---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rocnp

<!-- badges: start -->
[![R-CMD-check](https://github.com/dragosmg/rocnp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dragosmg/rocnp/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/dragosmg/rocnp/graph/badge.svg)](https://app.codecov.io/gh/dragosmg/rocnp)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/rocnp)](https://CRAN.R-project.org/package=rocnp)
<!-- badges: end -->

The goal of {rocnp} is to provide a set of functions for working with Romanian Personal Numeric Codes / Coduri Numerice Personale (CNP). 

## Features

rocnp includes the following functionality for working with Romanian personal numeric codes (PNC / CNP):

* an S3 class called `cnp`:
  * implemented as a record (similar to `POSIXlt`) with `vctrs::new_rcrd()`. The implementation details are 
* a constructor, `cnp()` for creating a `cnp` object 
* the constructor automatically decomposes the CNP and augments it by parsing the various field
* access the various components with the `extract_()` family of functions:
    * use `extract_sex()` for sex.
    * `extract_birth_year()` for the year of birth.
    * `extract_birth_month()` for the month of birth.
    * `extract_county()` for county of issue.
    * `extract_status()` for the residence status.
    * `extract_dob()` for the date of birth.

## Installation

You can install the released version of {rocnp} from [CRAN](https://CRAN.R-project.org):

``` r
install.packages("rocnp")
```

Alternatively, if you need the development version from [GitHub](https://github.com/dragosmg/rocnp), install it with

``` r
# install.packages("pak")
pak::pkg_install("dragosmg/rocnp")
```

## Usage

This is an example which shows you how to use some of the functions in rocnp:

```{r example}
library(rocnp)

# these are synthetically generated CNPs
# check CNP is valid
cnps <- cnp(
  c(
    "1940616346114", "6201206018078", "1940616346114"
  )
)

extract_sex(cnps)
extract_birth_year(cnps)
extract_birth_month(cnps)
extract_dob(cnps)
extract_county(cnps)
extract_status(cnps)
```

